Trafficking synaptic cargo involves tradeoffs in speed and precision

### 2015-12-09 latex/paper-v5.tex

\documentclass[11pt]{wlpeerj}
\usepackage{mathtools}
\usepackage{soul}
\usepackage{setspace}
\doublespacing

@@

\begin{abstract}
Nervous system function depends on the appropriate development, maintenance, and plasticity of synaptic connectivity. In order to form and modify synaptic connections, organelles and molecular cargo need to be trafficked from the site of synthesis to the vicinity of the synapse.
An open question is how synaptic cargo can be delivered to appropriate synapses in spite of the complex dendritic arbors and thousands of postsynaptic sites typical of many neurons.
We develop a biophysically-rooted yet general model of microtubule transport that captures known mechanistic properties of intracellular transport. In spite of the assumption that molecular transport responds only to local chemical cues, we show that arbitrarily complex distributions of cargo can be achieved. However, the model reveals important tradeoffs that can be tested experimentally.
We find that precise and flexible delivery of synaptic cargo requires the capture of synaptic cargo to occur on a slower timescale than the movement of cargo across the microtubule network.
Tuning the transport parameters reveals that optimal solutions exist for specific distributions of synapses, but these solutions sensitively depend on the location as well as the number of synapses, suggesting that biological neurons face a compromise between flexibility of connectivity and efficacy in regulating synaptic function.

@@

Neurons often have complex dendritic trees, reflecting the dense and structured synaptic connectivity that supports nervous system function.
In order to maintain appropriate connectivity, or to alter it in response to changes in circuit activity, numerous enzymes and macromolecules must be repeatedly replenished and redistributed amongst a staggering number of synaptic contacts.
Although the molecular machinery that solves this logistical problem is known in increasing detail [REFS], there are few attempts to quantify broadly how subcellular transport processes might constrain a neuron's ability to regulate its connectivity and synaptic function.

Transport along dendrites is mediated by motor proteins (kinesins and dyneins) that walk along microtubules, carrying cargo at a rate that far exceeds passive diffusion \citep{Smith_2001}\hl{Others?}.
The mechanisms that detach and sequester cargo at specific places in dendrites are not fully understood, but are known to depend on local biochemical signals such as transient elevations in cyclic AMP or $[Ca^{2+}]$ \citep{Mironov_2007,Wang_2009}, that are caused by synaptic activity or locally released growth factors from other cells.
In this way, the proteins, scaffolding molecules and receptors that make up a postsynaptic site can be made available to alter an existing synapse or assemble a new synapse in the vicinity of a presynaptic partner.
Not all aspects of synaptic plasticity require delivery of cargo from the soma. For example, some forms of long-term potentiation (LTP) depend on local protein synthesis and can function in isolated dendrites [REFS].
Nonetheless, components originating in the soma need to find their way to synapses in the first place, and in the long run, the mRNAs and the machinery that supports local biosynthesis needs to be replenished.
Moreover, many long-lasting forms of synaptic plasticity are known to depend on anterograde transport of mRNAs \citep{Kandel_2001,Puthanveettil_2008} and specific mRNAs are known to be selectively transported to regions of heightened synaptic activity \citep{Steward_1998,Steward_2001,Moga_2004} or to developing synaptic contacts \citep{Lyles_2006}.
These observations fit with the well-known synaptic tagging hypothesis \citep{Frey_1997}, which proposes that synapses produce a biochemical `tag' that signals a requirement for synaptic building blocks as part of the plasticity process.

The distribution of molecular cargo is partially controlled by a coarse addressing system in the localization sequences of 3$'$-untranslated regions of mRNAs -- which, for example, control whether mRNA is sequestered in the nucleus or into dendrites \citep{Andreassi2009}.
However, it is unclear whether a fine-scale, centralized addressing system, as exemplified by human-designed transport systems, is feasible or even desirable for neurons to implement.
Other biological systems (e.g. ant colonies) utilize decentralized transport strategies, which emerge from simple, local decisions made by a distributed population of agents \citep{Deneubourg_1990}.
Such self-organizing systems often enjoy a number of advantageous properties -- including scalability, flexibility, and fault-tolerance -- and are therefore a current topic of substantial interest to engineers \citep{Werfel2014,Rubenstein2014}.

%From simple, experimentally verifiable assumptions about the dynamics of active dendritic transport we develop a minimal biophysical model that can account for activity-dependent accumulation of molecular cargo at synaptic sites in realistic neuronal morphologies.
We develop a minimal biophysical model that can account for activity-dependent accumulation of molecular cargo at synaptic sites in realistic neuronal morphologies, and show how local biochemical signals can modify the movement and/or release of cargo to produce complex spatial patterns in a self-organizing fashion.
However, the decentralized transport strategies we explore face unavoidable tradeoffs between the speed and precision of cargo delivery.
Tuning the transport parameters reveals that while optimal solutions exist to balance this tradeoff, these solutions sensitively depend on the location as well as the number of synapses, suggesting that biological neurons must compromise between flexibility in connectivity and their ability to precisely regulate synaptic function.

@@

\hl{Activation of synaptic receptors triggers downstream signalling cascades that can alter global gene expression [ref Greenberg, Bading, others] as well as modulating the trafficking and metabolism of molecular cargo locally [REFS]. [more on this] \textbf{move to intro?}}

Transport along microtubules is mediated by kinesin and dynein motors, which mediate anterograde and retrograde transport, respectively \citep{Hirokawa_2010,Gagnon_2011}.
Cargo is often simultaneously bound to both forms of motor protein, resulting in stochastic back-and-forth movements with a net direction determined by the balance of opposing movements \citep{Hancock_2014,Buxbaum_2014b} (Fig. 1A).

To obtain a general model that can accommodate variations in the biophysical details, we consider microtubule-based transport as a biased random walk along a one-dimensional cable \citep{Bressloff_2006,Bressloff_2009,Newby_2010}.
For each time step, the cargo moves forwards, backwards, or remains in the same place with different probabilities that together sum to 1.
In the simplest model, the probabilities associated with each movement are fixed and independent across each time step, with the forward jump more probable than a reverse jump, leading to a biased random walk (Fig. 1B, top panel).
A potentially more realistic model (see, e.g., \cite{Soundararajan_2014}) incorporates hysteresis into the biased random walk (see Methods). In this case, the cargo is more likely to continue in the same direction as it moved during the previous time step, leading to extended runs in one direction along the microtubule (Fig. 1B, bottom panel).

\begin{figure}[!t]
\begin{center}
\includegraphics[width=0.9\columnwidth]{00_stochastic.png}
\caption{Model of microtubular transport.
\textbf{(A)} Synaptic cargo, $u$, undergoes stochastic back-and-forth movements driven by opposing motor proteins.
\textbf{(B)} Three example random walks on a cable representing the movement of individual cargoes. In a simple random walk, each movement is independent of previous movements (top panel); longer run lengths result from adding history-dependence to the model, such that each movement is likely to continue in the same direction as the previous step (bottom panel).
\textbf{(C)} An ensemble of synaptic cargoes transported along the length of a neurite.
\textbf{(D)} The concentration profile of transported cargo along a cable over time, simulated as 1,000 independent random walks. Simulations with (bottom) and without (top) history-dependence.
\textbf{(E)} In the limit of many individual particles, the concentration of $u$ is modeled by the convection-diffusion partial differential equation. The parameters, $a$ and $b$, respectively scale the anterograde and retrograde rate of transport (see equation 1).
\textbf{(F)} The convection-diffusion model provides a good fit for the simulations in panel D.
}
\end{center}
\end{figure}

While the position of individual cargoes can be highly stochastic, the net movement of a population of cargoes (Fig. 1C) is predictable.
Figure 1D shows the distribution of 1000 molecules over time with (top panel) and without (bottom panel) hysteresis.
For the simple biased random walk and for the random walk with hysteresis, these simulations can be well-fit by an advection-diffusion equation (Fig. 1E-F), which is easier to analyze because it is deterministic and contains fewer variables \citep{Smith_2001}.

We model transport using the advection-diffusion equation with spatially non-uniform drift and diffusion coefficients. As an approximation, we discretized a dendritic tree into small compartments, and modeled the movement of cargo between neighboring compartments as reactions with first order kinetics. In an unbranched cable with $N$ compartments, the mass-action model is:
\begin{equation}
u_1 \xrightleftharpoons[b_1]{a_1} u_2 \xrightleftharpoons[b_2]{a_2} u_3 \xrightleftharpoons[b_3]{a_3} ~...~ \xrightleftharpoons[b_{N-1}]{a_{N-1}} u_N
\end{equation}
where $u_i$ is the amount of cargo in each compartment, and $a_i$ and $b_i$ respectively denote local rate constants for anterograde and retrograde transport; these parameters determine the local diffusion and drift coefficients as shown in figure 1E (see Methods). Similar mass-action schemes can be constructed to model transport in branched morphologies (Fig. 2A). The mass-action scheme provides a flexible way to model local changes in transport rates, and is very tractable to mathematical analysis and numerical simulation.

A central assumption of mass-action description of this model is that the net movement of molecules between neighboring compartments is a stochastic and memoryless process. This assumption is reasonable for cargoes that change direction often; specifically, on a length scale comparable to the size of the compartments. This appears to be the case under many \citep{Muller_2008,Verbrugge_2009}, though not necessarily all \citep{Dynes_2007,Soundararajan_2014} circumstances. In simulations, we observed that this model was not very sensitive to violations of this assumption; the population dynamics of cargoes with substantial run lengths could still be well-fit by this class of models (Fig. 1F, lower panel; Supplementary Fig?).

@@

We next show how the mass-action model can be easily solved to produce complex distributions in the concentration of cargo along a dendrite
The exchange of $u$ between neighboring compartments approaches an equilibrium/steady-state ($ss$) distribution over time. In an unbranched dendrite, this occurs precisely when:
\begin{equation}
\frac{u_i}{u_{i+1}} \Bigg|_{ss} = \frac{b_i}{a_i}
\end{equation}
(see Methods). Thus, the local trafficking rates, $a_i$ and $b_i$, can be tuned to produce a desired spatial distribution of cargo. Let $\tilde{u}$ denote the desired steady-state level of $u$ in each compartment. For example, $\tilde{u}_i$ might refer to the local concentration of biochemical signal that ``tags'' synapses for plasticity \citep{Frey_1997}; such a tag might be positively correlated to the desired level of delivered cargo. To match this desired distribution, the ratio of local trafficking rates should satisfy:
\begin{equation}
\frac{b_i}{a_i} = \frac{\tilde{u}_i}{\tilde{u}_{i+1}}
\end{equation}

\begin{figure}[!tb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{01_mass_action.png}
\caption{Local trafficking parameters determine the spatial distribution of biomolecules by a simple rule.
\textbf{(A)} Diagram of the mass action transport model for a simple branched morphology.
%The steady-state ratio of $u$ between any two neighboring compartments is equal to the reciprocal ratio of the trafficking rate constants between those compartments.
\textbf{(B)} A simulation of a nine compartment cable, with trafficking rate constants set to produce a linear gradient using the steady-state relation shown in panel A.
%The initial distribution of $u$ is uniform (t=0); over time the system achieves the desired linear gradient (steady-state occurs around t=5, in this example).
\textbf{(C)} The slope of the linear gradient shown in panel B is determined by the total amount of $u$ present in the model; the slope increases as more of $u$ is present, but the linear profile is preserved.
\textbf{(D)} A model of a CA1 pyramidal cell with 742 compartments based on \cite{Migliore_2012}; excitatory synapses were added at the locations marked by red dots. 
\textbf{(E)} The average membrane potential in each compartment of the CA1 model cell was used to define a desired spatial distribution, and transport rate constants were chosen according to equation (4). The desired spatial distribution of cargo emerges over time, after starting in the soma.
\textbf{(F)} The desired and steady-state profile of $u$ for each compartment in the simulation shown in panel E.
}
\end{center}
\end{figure}

For example, a linear expression pattern emerges when $\tilde{u}_i \propto i$, resulting in $b_i / a_i = i / i + 1$ (Fig. 2B).
This rule produces the expected linear expression gradient with the slope controlled by tuning the total amount of cargo in the cable (Fig. 2C).

These results extend directly to branched morphologies (see Methods). For any pair of connected compartments:
\begin{equation}
%\frac{u_p}{u_c} \Bigg|_{ss} = \frac{b}{a}
\frac{b}{a} = \frac{\tilde{u}_p}{\tilde{u}_c}
\end{equation}
where $u_p$ is the level in the ``parent" compartment (closer to soma) and $u_c$ is the level in the ``child" compartment (closer to periphery); $b$ and $a$ refer to the local transport rate constants between this pair of compartments.

Together, this analysis shows that by slowing or biasing local trafficking rates, arbitrary distributions of cargo can be achieved over time. Local trafficking rates, in turn, can be modulated by a simple biochemical reaction resulting from synaptic activation or local release of a growth factor. We can consider the local cytoplasmic calcium concentration, $[ca]_i$, as an candidate second messenger. Increases in $[ca]_i$ simultaneously arrest anterograde and retrograde microtubular transport \citep{Wang_2009}. That is, for any pair of compartments, the anterograde rate constant is determined by calcium in the parent compartment, $a = f([ca]_p)$,  and the retrograde rate constant is determined by calcium in the child compartment, $b = f([ca]_c)$, where $f$ is a function that describes how calcium alters the transport rates. This results in:
\begin{equation}
\frac{b}{a} = \frac{f([ca]_c)}{f([ca]_p)} = \frac{\tilde{u}_p}{\tilde{u}_c}
\end{equation}
where $\tilde{u}_i = 1/f([ca]_i)$.

Thus, in principle, local calcium transients that serve to arrest transport could provide a mechanism for matching cargo concentration to an arbitrarily complicated distirbution over a dendritic tree.
To test the plausibility of this model in a complex morphology, we extended an existing multi-compartment model of a CA1 pyramidal neuron \citep{Migliore_2012}. Excitatory synaptic input was delivered to 120 locations within three dendritic regions (red dots, Fig. 2D), and the average membrane potential in each electrical compartment determined the target level $\tilde{u}_i$ in each compartment (Methods). This models how molecular cargo could be selectively trafficked to active synaptic sites. Figures 2E and 2F confirm that the spatial distribution of $u$ approaches the desired steady-state.

@@

Biochemical processes are time-sensitive.
For example, newly synthesized proteins must be delivered to synapses within $\sim$1 hour to support long-term potentiation in CA1 pyramidal cells \citep{Frey_1997,Frey_1998}.
It is therefore important to know what factors influence the rate at which the distribution of cargo matches its demand profile across the whole dendritic tree.

In equations (4) and (5), we implicitly required each $\tilde{u}_i > 0$ in order to avoid division by zero.
Intuitively, if $\tilde{u}_i = 0$ for some compartment, then no cargo can flow through that compartment, cutting of more peripheral compartments from the transport system.
Similarly, if certain $\tilde{u}_i$ are nearly equal to zero, then transport through these compartments will act as a bottleneck for transport, and convergence to the desired distribution will be slow.

Figure 3 illustrates and analyzes a bottleneck in a simple three compartment model.
The two compartments on the end of the cable have the same desired level, $\tilde{u}_1 = \tilde{u}_3$; the middle compartment acts as a bottleneck when $\tilde{u}_2$ is very small (Fig. 3A).
This results in the mass-action model:
\begin{equation}
u_1 \underset{x}{\overset{\epsilon}{\rightleftharpoons}} u_2 \underset{\epsilon}{\overset{x}{\rightleftharpoons}} u_3
\end{equation}
We assume that $u$ begins at one end of the cable, and examine the how taking $\epsilon$ to zero affects the convergence rate.
Figure 3B shows that the convergence rate slows dramatically as $\epsilon$ decreases.
This can be can be solved by analysing the corresponding system of ordinary differential equations. The convergence rate is determined by the non-zero eigenvalue with smallest magnitude, which can be thought of as a rate-limiting step or process for the system (Methods, Supp. Fig. 1).
Simulations confirmed this analysis and showed that the convergence rate diverges to infinity as $\epsilon$ approaches zero (Fig. 3C).
\hl{I think the epsilon goes to zero stuff can be omitted, or at the very least stated more tersely - if the transport rate is zero anywhere it is obvious that it takes forever to converge}

\begin{figure}[!tb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{03_convergence.png}
\caption{ Convergence to steady-state is slow when molecules must be transported across bottlenecks ---compartments with low target levels.
\textbf{(A)} A three-compartment transport model, with the middle compartment acting as a bottleneck. The vertical bars represent the desired steady-state level of $u$ in each compartment. The rate of transport into the middle compartment is small ($\epsilon$, dashed harpoons) relative to transport out of the middle compartment.
\textbf{(B)} As $\epsilon$ decreases, the model converges more slowly and the steady-state level decreases in the middle compartment.
\textbf{(C)} Simulations (black dots) confirm that the time to convergence is given by the smallest non-zero eigenvalue of the system (analytically calculated line). This eigenvalue can be thought of as the rate-limiting step of the system.
\textbf{(D)} Convergence ($L_1$ distance) to a uniform target distribution (red line) is faster than a target distribution containing a bottleneck (blue line) in the CA1 model.
\textbf{(E)} For all compartments that reach a threshold level ($u_i = 0.001$), the simulated time it takes to reach threshold is plotted against the distance of that compartment to the soma.
\textbf{(F)} Predicted convergence times for various target distributions (str. rad., stratum radiatum; str. lac., stratum lacunosum/moleculare).}
\end{center}
\end{figure}

Qualitatively similar results were obtained in the multi-compartment CA1 neuron model.
The model converged to a uniform target distribution more quickly than to a ``bottleneck" target distribution, in which the middle third of the apical dendrite had low steady-state levels of cargo (Fig. 3D).
Each pair of anterograde and retrograde rate constants were normalized to sum to one; thus, differences in convergence were not due to the scale of the trafficking rate constants.

>>>>

\documentclass[11pt]{wlpeerj}
\usepackage{mathtools}
\usepackage{soul}
\usepackage{setspace}
%\doublespacing

@@

\begin{abstract}
Neurons form, maintain, and modify complex connectivity patterns by regulating the spatial distribution of organelles and macromolecules throughout their dendritic arbors. Previous work suggests that localized biochemical signals influence the trafficking, delivery and expression of molecular cargo along microtubules. However, it remains unclear whether purely local mechanisms can be orchestrated to provide adequate transport. We mathematically formalized a previously proposed ``sushi belt'' model of microtubule transport \citep{Doyle_2011} and showed that arbitrarily complex spatial distributions of cargo can indeed be achieved by localized mechanisms. However, this model predicts an unavoidable tradeoff between speed and precision of cargo delivery that can be tested experimentally. Reliable cargo delivery required that cargo transport along microtubules occur on a faster timescale than cargo dissociation from the microtubules. Increasing the cargo dissociation rate provided faster transport, but required complex, global re-tuning of the trafficking rates to maintain precision. Additionally, these solutions tended to be fragile in the sense that small changes in the target distribution led to large changes in the optimal delivery strategy. Together, these results suggest that macromolecules with stereotyped expression patterns might be tuned for fast and precise transport, but cargoes requiring plastic expression patterns face a tradeoff in the speed and precision of transport.

@@

Neurons often have complex dendritic trees, reflecting the dense and structured synaptic connectivity that supports nervous system function.
In order to maintain appropriate connectivity, or to alter it in response to changes in circuit activity, numerous enzymes and macromolecules must be continually replenished and distributed among a large number of synaptic contacts.
Although the molecular machinery that solves this logistical problem has been characterized in increasing detail \citep{Doyle_2011,Buxbaum_2014b,Hancock_2014}, there are few attempts to formalize how subcellular transport processes constrain a neuron's ability to regulate its connectivity and synaptic function.

Not all aspects of neurite metabolism require delivery of cargo from the soma. For example, some forms of long-term potentiation (LTP) depend on local protein synthesis and can function in isolated dendrites \citep{Kang1996,Aakalu:2001aa,vickers2005induction,Sutton:2006aa}.
Nonetheless, components originating in the soma need to find their way to synapses in the first place, and in the long run, the mRNAs and the machinery that supports local biosynthesis needs to be replenished.
Moreover, many long-lasting forms of synaptic plasticity are known to depend on anterograde transport of mRNAs \citep{Kandel_2001,Puthanveettil_2008} and specific mRNAs are known to be selectively transported to regions of heightened synaptic activity \citep{Steward_1998,Steward_2001,Moga_2004} or to developing synaptic contacts \citep{Lyles_2006}.
These observations fit with the well-known synaptic tagging hypothesis \citep{Frey_1997}, which proposes that synapses produce a biochemical `tag' that signals a requirement for synaptic building blocks as part of the plasticity process.

Kinesin and dyenin motor proteins mediate transport along microtubules at rates that far exceed passive diffusion \citep{Block_1990,Smith_2001,Hirokawa_2010,Gagnon_2011,park2014visualization}.
In many cases, the movement of individual cargoes is stochastic and bidirectional \citep{Hancock_2014}, leading to the hypothesis that motor proteins ``search'' the dendritic tree for dropoff locations, instead of following a pre-determined path \citep{Welte2004,Bressloff_2009,Doyle_2011}.
Thus, the local rate of transport and the decision to release cargo are thought to be controlled by localized signaling pathways, rather than a centralized addressing system.
These localized mechanisms are not fully understood, but are believed to involve transient elevations in second-messengers like $[Ca^{2+}]$ and ADP \citep{Mironov_2007,Wang_2009,Buxbaum_2014a}, and changes in microtubule-associate proteins \citep{Soundararajan_2014}.
These pathways are, in turn, influenced by the spatial pattern of synaptic activity, spontaneous release of calcium from internal stores, and locally released factors from other cells \citep{Wong2002,Parrish2007,Park2013,Zagrebelsky2014}.
The composition and expression of molecular cargo is also influenced by global factors, such as the cell's average activity level \citep{Spitzer2006,Puthanveettil_2008,Turrigiano2011} and localization sequences within 3$'$-untranslated regions of mRNA, which act as a coarse addressing system \citep{Andreassi2009}.

%However, it is not know how intracellular cargo can be targeted on a finer scale in a way that responds to varying demand.

Here, we aim to express and examine existing conceptual models of molecular transport \citep{Doyle_2011} in a mathematical framework.
We develop a minimal biophysical model that can account for activity-dependent accumulation of molecular cargo at synaptic sites in realistic neuronal morphologies, and show that biochemical signals can indeed produce complex spatial expression patterns by locally modifying the movement and release of cargo.
However, we find that decentralized transport strategies face unavoidable tradeoffs between the speed and precision of cargo delivery.
Tuning the transport parameters reveals that while optimal solutions exist to circumvent this tradeoff, these solutions sensitively depend on the location as well as the number of synapses, suggesting that biological neurons may face a compromise between flexibility in connectivity and their ability to precisely regulate synaptic function.

@@

Transport along microtubules is mediated by kinesin and dynein motors, which mediate anterograde and retrograde transport, respectively \citep{Hirokawa_2010,Gagnon_2011}.
Cargo is often simultaneously bound to both forms of motor protein, resulting in stochastic back-and-forth movements with a net direction determined by the balance of opposing movements \citep{Welte2004,Hancock_2014,Buxbaum_2014b} (Fig. 1A).

To obtain a general model that can accommodate variations in the biophysical details, we consider microtubule-based transport as a biased random walk along a one-dimensional cable \citep{Bressloff_2006,Bressloff_2009,Newby_2010,Bressloff_2013}.
For each time step, the cargo moves forwards, backwards, or remains in the same place with different probabilities that together sum to 1.
In the simplest model, the probabilities associated with each movement are fixed and independent across each time step, with the forward jump more probable than a reverse jump, leading to a biased random walk (Fig. 1B, top panel).
A potentially more realistic model (see, e.g., \cite{Muller_2008}) incorporates hysteresis into the biased random walk (see Methods). In this case, the cargo is more likely to continue in the same direction as it moved during the previous time step, leading to extended runs in one direction along the microtubule (Fig. 1B, bottom panel).

\begin{figure}[!t]
\begin{center}
\includegraphics[width=0.9\columnwidth]{00_stochastic.png}
\caption{Model of microtubular transport.
\textbf{(A)} Molecular cargo, $u$, undergoes stochastic back-and-forth movements driven by opposing motor proteins.
\textbf{(B)} Three example random walks on a cable representing the movement of individual cargoes. In a simple random walk, each movement is independent of previous movements (top panel); longer run lengths result from adding history-dependence to the model, such that each movement is likely to continue in the same direction as the previous step (bottom panel).
\textbf{(C)} An ensemble of synaptic cargoes transported along the length of a neurite.
\textbf{(D)} The concentration profile of transported cargo along a cable over time, simulated as 1,000 independent random walks. Simulations with (bottom) and without (top) history-dependence.
\textbf{(E)} In the limit of many individual particles, the concentration of $u$ is described by a mass-action model (equation 1). The parameters $a$ and $b$ respectively scale the anterograde and retrograde rate of transport.
\textbf{(F)} The mass-action model provides a good fit for the simulations in panel D.
}
\end{center}
\end{figure}

While the position of individual cargoes is highly stochastic, the net movement of a population of cargoes (Fig. 1C) is predictable.
Figure 1D shows the distribution of 1000 molecules over time with (top panel) and without (bottom panel) hysteresis.
Thus, the bulk trafficking of cargo can be modeled as a deterministic process, which we refer to as the ``mass-action model'' of transport. The model discretizes the dendritic tree into small compartments, and describes the transfer of cargo between neighboring compartments as reactions with first order kinetics. In a cable with $N$ compartments, the mass-action model is (Fig. 1E):
\begin{equation}
u_1 \xrightleftharpoons[b_1]{a_1} u_2 \xrightleftharpoons[b_2]{a_2} u_3 \xrightleftharpoons[b_3]{a_3} ~...~ \xrightleftharpoons[b_{N-1}]{a_{N-1}} u_N
\end{equation}
where $u_i$ is the amount of cargo in each compartment, and $a_i$ and $b_i$ respectively denote local rate constants for anterograde and retrograde transport. The mass-action model with spatially uniform transport rates (Fig. 1F) reproduces the dynamics of the ensemble simulations (Fig. 1D). In this case, the mass-action model maps onto the well-known drift-diffusion equation (see \textit{Methods}).

The mass-action model makes two important assumptions. It first assumes that the net movement of molecules between neighboring compartments is a stochastic and memoryless process. This assumption is reasonable for cargoes that change direction often; specifically, on a length scale comparable to the size of the compartments. This appears to be the case under many \citep{Muller_2008,Verbrugge_2009}, though not necessarily all \citep{Dynes_2007,Soundararajan_2014} circumstances. However, we observed that the mass-action model could still be well-fit to simulated data where this assumption was violated (Fig. 1F, lower panel). In fact, the mass-action model only breaks down in the limit of very large run lengths (i.e. unidirectional, non-stochastic transport; see \textit{Methods}).

The second assumption of the mass-action model is that there are many transported particles distributed throughout the dendritic tree. Many types of dendritic cargo are present in high numbers \hl{(CITE)}, and a deterministic model can provide a good approximation in this regime (Fig. 1D). The model also provides insight into the stochastic dynamics of transport for cargoes with fewer copy numbers. Instead of interpreting $u$ as the amount of cargo in each compartment, this variable (when appropriately normalized) can be interpreted as the probability that a single particle lies within a particular compartment. The mass-action model captures average, or expected, behavior of the ensemble, with noise inversely proportional to the square root of the number of copies of cargo (see \textit{Methods}).

@@

We used the mass-action model for our simulations because it is easy to extend it to branched morphologies (Fig. 2A), and to cases where the trafficking rate constants ($a_i$ and $b_i$) are spatially non-uniform.
This corresponds to the kinetics of motor proteins depending on local biochemical signals \citep{Mironov_2007,Wang_2009,Soundararajan_2014}.

In the model, the exchange of cargo between compartments approaches a equilibrium, or steady-state ($ss$), distribution over time (see \textit{Methods}). The equilibrium occurs when no net movement of cargo occurs between connected compartments. Mathematically, this occurs when:
\begin{equation}
\frac{u_p}{u_c} \Bigg|_{ss} = \frac{b}{a}
%\frac{b}{a} = \frac{\tilde{u}_p}{\tilde{u}_c}
\end{equation}
where $u_p$ is the level in the ``parent" compartment (closer to soma) and $u_c$ is the level in the ``child" compartment (closer to periphery); $b$ and $a$ refer to the trafficking rate constants between this pair of compartments.

Intuitively, in the mass-action model, the rate of cargo transfer is proportional to the concentration of cargo and the trafficking rate constant. Thus, the flow of cargo is equal and opposite when the ratio of cargo matches the reciprocal ratio of the rate constants between a pair of connected compartments.

The above result means that the trafficking rates can be locally tuned to produce a desired spatial distribution of cargo. Denote the desired steady-state level of cargo in each compartment by $\tilde{u}$ (we will soon show how this can be encoded by a localized biochemical signal). To match this desired distribution, the ratio of local trafficking rates should satisfy:
\begin{equation}
\frac{b}{a} = \frac{\tilde{u}_p}{\tilde{u}_c}
\end{equation}

\begin{figure}[!tb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{01_mass_action.png}
\caption{Local trafficking parameters determine the spatial distribution of biomolecules by a simple rule.
\textbf{(A)} Diagram of the mass action transport model for a simple branched morphology.
%The steady-state ratio of $u$ between any two neighboring compartments is equal to the reciprocal ratio of the trafficking rate constants between those compartments.
\textbf{(B)} A simulation of a nine compartment cable, with trafficking rate constants set to produce a linear gradient using the steady-state relation shown in panel A.
%The initial distribution of $u$ is uniform (t=0); over time the system achieves the desired linear gradient (steady-state occurs around t=5, in this example).
\textbf{(C)} The slope of the linear gradient shown in panel B is directly proportional to the total amount of cargo in the model ($\sum_i u_i$); the slope increases with increasing cargo levels, but the linear profile is preserved.
\textbf{(D)} A model of a CA1 pyramidal cell with 742 compartments adapted from \cite{Migliore_2012}; excitatory synapses were added at the locations marked by red dots. 
\textbf{(E)} The average membrane potential in each compartment of the CA1 model cell was used to define a desired spatial distribution, and transport rate constants were chosen according to equation (4). The desired spatial distribution of cargo emerges over time, after starting in the soma.
\textbf{(F)} The desired and steady-state profile of cargo for each compartment in the simulation shown in panel E.
}
\end{center}
\end{figure}

For example, we can produce a linear expression gradient \citep{Magee_1998,Hoffman_1997} by setting $\tilde{u}$ directly proportional to the distance from the soma.
This produces trafficking rate constants that satisfy $b_i / a_i = i / i + 1$ (where $i$ indexes on increasing distance to the soma).
Figure 2B-C shows that this rule produces the expected profile, with the slope of the linear gradient controlled by tuning the total amount of cargo in the cable (Fig. 2C).

Together, this analysis shows that by manipulating local trafficking rates, arbitrary distributions of cargo can be achieved over time. Local trafficking rates, in turn, can be modulated by a simple biochemical reaction resulting from synaptic activation or local release of a growth factor. We can consider the local cytoplasmic calcium concentration, $[ca]_i$, as an candidate second messenger. Increases in $[ca]_i$ simultaneously arrest anterograde and retrograde microtubular transport for certain cargoes \citep{Wang_2009}. Thus, for any pair of compartments, the anterograde rate constant is determined by calcium in the parent compartment, $a = f([ca]_p)$,  and the retrograde rate constant is determined by calcium in the child compartment, $b = f([ca]_c)$, where $f$ is a function that describes how calcium alters the transport rates. This results in:
\begin{equation}
\frac{b}{a} = \frac{f([ca]_c)}{f([ca]_p)} = \frac{\tilde{u}_p}{\tilde{u}_c}
\end{equation}
where $\tilde{u}_i = 1/f([ca]_i)$.

Thus, in principle, local calcium transients that serve to arrest transport could provide a mechanism for distributing cargo to an arbitrary target profile.
To demonstrate the plausibility of this model in a complex morphology, we extended an existing multi-compartment model of a CA1 pyramidal neuron \citep{Migliore_2012}. Excitatory synaptic input was delivered to 120 locations within three dendritic regions (red dots, Fig. 2D), and the average membrane potential in each electrical compartment determined the target level ($\tilde{u}_i$) in each compartment (Methods). This models how molecular cargo could be selectively trafficked to active synaptic sites. Figures 2E and 2F confirm that the spatial distribution of $u$ approaches the desired steady-state.

@@

Biochemical processes are time-sensitive.
For example, newly synthesized proteins must be delivered to synapses within $\sim$1 hour to support long-term potentiation in CA1 pyramidal cells \citep{Frey_1997,Frey_1998}.
We therefore examined how quickly the spatial pattern of cargo converged to its target distribution.

In equations (3) and (4), we implicitly required each $\tilde{u}_i > 0$ in order to avoid division by zero.
Intuitively, if $\tilde{u}_i = 0$ for some compartment, then no cargo can flow through that compartment, cutting of more peripheral compartments from the transport system.
Similarly, if certain $\tilde{u}_i$ are nearly equal to zero, then transport through these compartments will act as a bottleneck for transport, and convergence to the desired distribution will be slow.

Figure 3A-C illustrates and analyzes a bottleneck in a simple three compartment model.
The two compartments on the end of the cable have the same desired level, $\tilde{u}_1 = \tilde{u}_3$; the middle compartment acts as a bottleneck when $\tilde{u}_2$ is very small (Fig. 3A).
This results in the mass-action model:
\begin{equation}
u_1 \underset{x}{\overset{\epsilon}{\rightleftharpoons}} u_2 \underset{\epsilon}{\overset{x}{\rightleftharpoons}} u_3
\end{equation}
We assume that $u$ begins at one end of the cable, and examine the how taking $\epsilon$ to zero affects the convergence rate.
Figure 3B shows that the convergence rate slows dramatically as $\epsilon$ decreases.
The convergence rate is determined by the smallest magnitude, non-zero eigenvalue of the system (see Methods), which can be thought of as a rate-limiting step or process for the system.
Simulations confirmed this analysis and showed that the convergence rate diverges to infinity as $\epsilon$ approaches zero (Fig. 3C).

\begin{figure}[!tb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{03_convergence.png}
\caption{ Convergence to steady-state is slow when molecules must be transported across bottlenecks ---compartments with low target levels.
\textbf{(A)} A three-compartment transport model, with the middle compartment acting as a bottleneck. The vertical bars represent the desired steady-state level of cargo in each compartment. The rate of transport into the middle compartment is small ($\epsilon$, dashed harpoons) relative to transport out of the middle compartment.
\textbf{(B)} As $\epsilon$ decreases, the model converges more slowly and the steady-state level decreases in the middle compartment.
\textbf{(C)} Simulations (black dots) confirm that the time to convergence is given by the smallest non-zero eigenvalue of the system (analytically calculated line). This eigenvalue can be thought of as the rate-limiting step of the system.
\textbf{(D)} Convergence ($L_1$ distance) to a uniform target distribution (red line) is faster than a target distribution containing a bottleneck (blue line) in the CA1 model.
\textbf{(E)} For all compartments that reach a threshold level ($u_i = 0.001$), the simulated time it takes to reach threshold is plotted against the distance of that compartment to the soma.
\textbf{(F)} Predicted convergence times for various target distributions (\textit{str. rad.}, stratum radiatum; \textit{str. lac.}, stratum lacunosum/moleculare).}
\end{center}
\end{figure}

We then asked whether the intuition from this three-compartment model extended to a cell with realistic morphology.
We obtained qualitatively similar results.
The CA1 model converged to a uniform target distribution more quickly than to a ``bottleneck" target distribution, in which the middle third of the apical dendrite had low steady-state levels of cargo (Fig. 3D).
Each pair of anterograde and retrograde rate constants were normalized to sum to one; thus, differences in convergence were not due to the scale of the trafficking rate constants.
